### 2011-12-14 AffineLieAlgebras/Thesis/5.tex
Translation on Affine.m

Infinite-dimensional Lie algebras also have growing number of applications in physics for example in conformal field theory and in a study of  integrable systems. But infinite-dimensional algebras are much harder to investigate and the number of available computer programs is much smaller in this case.

Affine Lie algebras \cite{kac1990idl} constitute important and tractable class of infinite-dimensional Lie algebras. They are constructed as central extensions of loop algebras of (semi-simple) finite-dimensional Lie algebras and appear naturally in a study of Wess-Zumino-Witten and coset models in conformal field theory \cite{Walton:1999xc}, \cite{difrancesco1997cft}, \cite{Goddard198588}, \cite{Dunbar:1992gh}.

The structure of affine Lie algebras allows to adapt for them computational algorithms created for finite-dimensional Lie algebras  \cite{Fuchs:1996dd}, \cite{gannon2001algorithms}, \cite{kass1990ala}. The book \cite{kass1990ala} with the tables of multiplicities and other computed characteristics of affine Lie algebras and representations was published in 1990. But we are not aware of software packages for popular computer algebra systems which can be used to extend these results.
We address this issue and present {\bf Affine.m} -- {\it Mathematica} package for computations in representation theory of affine and finite-dimensional Lie algebras.  We describe the features and limitations of the package in the present paper.  We also provide representation-theoretical background of implemented algorithms and present examples of computations relevant to physical applications.

The paper starts with an overview of Lie algebras and their representation theory (Sec. \ref{sec:theor-backgr}). Then we describe datastructures of {\bf Affine.m} used to present different objects related to Lie algebras and representations (Sec. \ref{sec:core-datastructures}) and discuss implemented algorithms (Sec. \ref{sec:comp-algor}). Next section consists of physically interesting examples (Sec. \ref{sec:examples}). The paper is concluded with the discussion of possible extensions and refinements (Sec. \ref{sec:conclusion}).

\section{Theoretical background}
\label{sec:theor-backgr}

In this section we remind necessary definitions and present formulae used in computations.

%% Define external border of irreducible representation here

We can use the Weyl character formula to obtain recurrent relations for weight multiplicities -- important tools for calculations \cite{il2010folded,kulish4sfa}.

For irreducible highest-weight modules recurrent relations have the following form
\begin{equation}
\label{eq:14}
m_{\xi }=-\sum_{w\in W\setminus e}\epsilon (w)m_{\xi
-\left( w(\rho )-\rho \right) }+\sum_{w\in W}\epsilon
(w)\delta _{\left( w(\mu +\rho )-\rho \right) ,\xi }.
\end{equation}
Formulae for Verma and generalized Verma modules differ only in the second term on the right-hand side. In case of Verma module it is just $\delta_{\xi,\mu}$. For generalized Verma module the summation in the second term on the right-hand side of \eqref{eq:14} is over the Weyl subgroup generated by reflections corresponding to roots $\{\alpha_{I}\}$.

Other recurrent formula can be obtained from a study of Casimir
elements action on irreducible highest weight modules
\cite{humphreys1997introduction}:
\begin{equation}
  \label{eq:15}
  m_{\lambda}=\frac{2}{(\mu+\rho)^{2}-(\lambda+\rho)^{2}}\sum_{\alpha\in \Delta^{+}}\sum_{k\geq 1} (\lambda+k\alpha,\alpha)m_{\lambda+k\alpha}.
\end{equation}
It is called the Freudenthal multiplicity formula.
Note that it is applicable only to irreducible modules.

We discuss the use of formulae \eqref{eq:14} and \eqref{eq:15} for computations in section \ref{sec:comp-algor}.

Now consider an algebra $\gf$ and a reductive subalgebra
$\af\subset \gf$. Simple roots $\beta_{i}$ of the subalgebra $\af$
can be presented as linear combinations of $\gf$-algebra roots
$\alpha_{j}$: $\beta_{i}=\sum_{j=1,\dots,r_{\gf}}k_{j}
\alpha_{j},\ j=1,\dots,r_{\af}$.

Each irreducible $\gf$-module is also an $\af$-module, although
$L^{\mu}_{\gf}$ is in general not irreducible as $\af$-module. 
It can be
decomposed into a direct sum of irreducible $\af$-modules:
\begin{equation}
  \label{eq:16}
  L^{\mu}_{\gf}=\bigoplus_{\nu}b^{\mu}_{\nu}L^{\nu}_{\af}
\end{equation}
Coefficients in such a decomposition are called branching
coefficients.

It is possible to calculate branching coefficients by constructing and successively  subtracting the submodules $L^{\nu}_{\af}$. 
This traditional approach has serious limitations especially in case of affine Lie algebras. We discuss them in the end of section \ref{sec:comp-algor}.

Now we describe an alternative approach which is based on recurrent properties of branching coefficients. 
But before we proceed to these recurrent relations we need several additional definitions.

For a subalgebra $\af\subset \gf$ we introduce the subalgebra
$\afb$. Consider the root subspace $\hf_{\perp \af}^{\ast }$
orthogonal to $\hf_{\af}$,
\begin{equation*}
\hf_{\perp \af}^{\ast }:=\left\{ \eta \in \hf^{\ast }|\forall
h\in \hf_{\af};\eta \left( h\right) =0\right\} ,
\end{equation*}
and the roots (correspondingly -- positive roots) of $\gf$ orthogonal
to roots of $\af$,
\begin{eqnarray}
\Delta _{\afb} &:&=\left\{ \beta \in \Delta _{\gf}|\forall
h\in \hf_{\af};\beta \left( h\right) =0\right\} ,
\label{delta a ort} \\
\Delta _{\afb}^{+} &:&=\left\{ \beta ^{+}\in \Delta _{\gf%
}^{+}|\forall h\in \hf_{\af};\beta ^{+}\left( h\right) =0\right\} .
\notag
\end{eqnarray}
Let $W_{\afb}$ be a subgroup of $W$ generated by
reflections $w_{\beta }$ with the roots $\beta \in \Delta _{\af_{\perp
}}^{+}$. The subsystem $\Delta _{\afb}$ determines a 
subalgebra $\afb$ with the Cartan subalgebra $\hf_{\af%
_{\perp }}$.

The Cartan subalgebra $\frak{h}$ can be decomposed in the following way:  $\frak{h}=\frak{\frak{h}_{\af}}\oplus
\frak{h}_{\afb}\oplus \frak{h}_{\perp }$

We also introduce the notations
\begin{eqnarray}
\widetilde{\frak{a}_{\perp }} :=\frak{a}_{\perp }\oplus \frak{h}_{\perp }
\qquad
\widetilde{\frak{a}} :=\frak{a}\oplus \frak{h}_{\perp }.
\end{eqnarray}

For $\af$ and $\afb$ we consider the
corresponding Weyl vectors, $\rho _{\af}$ and $\rho _{\af_{\perp
}} $ and compose the so called ''defects'' $\mathcal{D}_{\af}$ and $\mathcal{%
D}_{\afb}$ of the injection:
\begin{equation}
\mathcal{D}_{\af}:=\rho _{\af}-\pi _{\af}\rho , \qquad
\mathcal{D}_{\afb}:=\rho _{\afb}-\pi _{\af%
_{\perp }}\rho .  \label{defect-ort}
\end{equation}

%%!!! Define P^+
For $\mu \in P^{+}$ consider the linked weights $\left\{ \left(
w(\mu +\rho )-\rho \right) |w\in W\right\} $ and their projections
to
$h_{\afb}^{\ast }$ additionally shifted by the defect $-%
\mathcal{D}_{\afb}$:
\begin{equation*}
\mu _{\afb}\left( w\right) :=\pi _{\afb}\left[
w(\mu +\rho )-\rho \right] -\mathcal{D}_{\afb},\quad w\in W.
\end{equation*}
Among the weights $\left\{ \mu _{\af_{\perp
}}\left( w\right) |w\in W\right\} $ one can always choose those located in
the fundamental chamber $\overline{C_{\afb}}$. Let $U$ be the
set of representatives $u$ for the classes $W/W_{\afb}$ such
that

\begin{equation}
U:=\left\{ u\in W|\quad \mu _{\afb}\left( u\right) \in
\overline{C_{\afb}}\right\} \quad .  \label{U-def}
\end{equation}
Thus we can form the subsets:
\begin{equation}
\mu _{\widetilde{\mathfrak{a}}}\left( u\right) :=\pi _{\widetilde{%
\mathfrak{a}}}\left[ u(\mu +\rho )-\rho \right] +\mathcal{D}_{\af%
_{\perp }},\quad u\in U,  \label{mu-a}
\end{equation}
and
\begin{equation}
\mu _{\afb}\left( u\right) :=\pi _{\afb}\left[
u(\mu +\rho )-\rho \right] -\mathcal{D}_{\afb},\quad u\in U.
\label{mu-a-tilda}
\end{equation}

Notice that the subalgebra $\mathfrak{a}_{\bot}$ is regular by definition
since it is built on a subset of roots of the algebra $\mathfrak{g}$.

Denote by  $k_{\xi }^{\left( \mu \right) }$ signed branching coefficients. If $\xi\in \bar C_{\af}$ is in main Weyl chamber $k_{\xi}^{(\mu)}=b^{(\mu)}_{\xi}$ otherwise $k_{\xi}^{(\mu)}=\epsilon(w) b^{(\mu)}_{w (\xi+\rho_{\af})-\rho_{\af}}$ where $w\in W_{\af}$ is such that $w (\xi+\rho_{\af})-\rho_{\af}\in \bar C_{\af}$.

Now we can use the Weyl character formula to write a
recurrent relation \cite{2010arXiv1007.0318L} for signed branching
coefficients $k_{\xi }^{\left( \mu \right) }$ corresponding to an
injection $\af\hookrightarrow \gf$:
\begin{equation}
\begin{array}{c}
k_{\xi }^{\left( \mu \right) }=-\frac{1}{s\left( \gamma _{0}\right) }\left(
\sum_{u\in U}\epsilon (u)\;\dim \left( L_{\afb}^{\mu _{\af%
_{\perp }}\left( u\right) }\right) \delta _{\xi -\gamma _{0},\pi _{%
\widetilde{\af}}(u(\mu +\rho )-\rho )}+\right.  \\
\left. +\sum_{\gamma \in \Gamma _{\widetilde{\af}\rightarrow \gf%
}}s\left( \gamma +\gamma _{0}\right) k_{\xi +\gamma }^{\left( \mu \right)
}\right) .
\end{array}
\label{recurrent-rel}
\end{equation}
The recursion is governed by the set $\Gamma _{\af\rightarrow \gf}$ called the injection fan. The latter is defined by the
carrier set $\left\{ \xi \right\} _{\af\rightarrow \gf}$ for the
coefficient function $s(\xi )$
\begin{equation*}
\left\{ \xi \right\} _{\widetilde{\af}\rightarrow \gf}:=\left\{
\xi \in P_{\widetilde{\af}}|s(\xi )\neq 0\right\}
\end{equation*}
appearing in the expansion
\begin{equation}
\prod_{\alpha \in \Delta ^{+}\setminus \Delta _{\bot }^{+}}\left( 1-e^{-\pi
_{\widetilde{\af}}\alpha }\right) ^{\mathrm{mult}(\alpha )-\mathrm{mult}%
_{\af}(\pi _{\widetilde{\af}}\alpha )}=-\sum_{\gamma \in P_{%
\widetilde{\af}}}s(\gamma )e^{-\gamma };\quad
\end{equation}
The weights in $\left\{ \xi \right\} _{\widetilde{\af}\rightarrow \gf}$ are to be shifted by $\gamma _{0}$ -- the lowest vector in $\left\{ \xi
\right\} $ -- and the zero element is to be eliminated:
\begin{equation}
\Gamma _{\af\rightarrow \gf}=\left\{ \xi -\gamma
_{0}|\xi \in \left\{ \xi \right\} \right\} \setminus \left\{ 0\right\} .
\end{equation}
Formula (\ref{eq:14}) is a particular case of recurrent relation for branching coefficients (\ref{recurrent-rel}) in the case of Cartan subalgebra $\af=\hfg$.

If the root system of $\afb$ is generated by some subset of $\gf$
simple roots $\alpha_{1},\dots,\alpha_{r}$ then the recurrent
relation (\ref{recurrent-rel}) is connected with the generalized
Bernstein-Bernstein-Gelfand resolution for parabolic Verma modules
\cite{2011arXiv1102.1702L}.

Another particular case of this formula is connected with tensor
product decompositions. Consider the tensor product of two
irreducible $\gf$-modules $L^{\mu}\otimes L^{\nu}$. It is also a
$\gf$-module but not irreducible in general. So
\begin{equation}
  \label{eq:19}
  L^{\mu}\otimes L^{\nu}=\bigoplus_{\gamma} f^{\mu\nu}_{\gamma}L^{\gamma}
\end{equation}
The coefficients $f^{\mu\nu}_{\gamma}$ are called fusion
coefficients. The problem of computation of fusion coefficients is
equivalent to branching problem for the diagonal subalgebra
$\gf\subset \gf\oplus \gf$ (see \cite{LyakhovskyPostnova2011}). So our implementation of recurrent algorithm can be used to decompose tensor products (See Section \ref{sec:tens-prod-decomp}).

In the case of affine Lie algebras $\gf, \af$  the multiplicities
$m_{\nu}$ and the corresponding branching coefficients $b_{\nu}$
can be regarded as the coefficients in power series decomposition
of string and branching functions correspondingly:
\begin{align}
  \label{eq:21}
  &\sigma_{\nu}(q)=\sum_{n=0}^{\infty} m_{\nu-n\delta} q^n, \quad \nu=\sum_j c_j \omega_j,\quad c_j\geq 0\\
  & b_{\nu}(q)=\sum_{n=0}^{\infty} b_{\nu-n\delta} q^n,\quad  \nu=\sum_j c_j \omega_j, \quad c_j\geq 0
\end{align}
String and branching functions have  modular and analytic properties which are important for conformal field theory, especially in  coset models and the study of CFT on higher genus surfaces \cite{kac1988modular}, \cite{difrancesco1997cft}, \cite{Walton:1999xc}, \cite{walton1989conformal}.

{\bf Affine.m}  calculates weight multiplicities and branching coefficients for affine Lie algebras up to some finite grade. We present examples of computations in Sections \ref{sec:string-funct-affine}, \ref{sec:branch-funct-coset}. Now we proceed to the description of datastructures and algorithms implemented in {\bf Affine.m}.

\section{Core datastructures}
\label{sec:core-datastructures} Having introduced necessary
mathematical objects, problems and relations we now describe the
related datastructures of {\bf Affine.m}. Although {\it
Mathematica} is untyped language it is possible to create
structured objects and do type checks with patterns
\cite{shifrinmathematica}, \cite{maeder2000computer}.
\subsection{Weights}
\label{sec:weights}

Weights are represented  by two data-structures: \lstinline{finiteWeight} for finite-dimensional Lie algebras and \lstinline{affineWeight}  for affine.

Internally the finite weight is a \lstinline{List} with the
\lstinline{Head} \lstinline{finiteWeight}, its components are the
coordinates of the weight in the orthogonal Bourbaki  basis
\cite{bourbaki2002lie}.

Affine weight is an extension of a finite weight by supplying it
with level and grade coordinates. There is a set of functions
defined for finite and affine weights. The complete list can be
found in  online help of the package. The most important
are definitions of addition, multiplication by a number and a
scalar product (bilinear form) for weights. These definitions
allow to use traditional notation when working with {\bf
Affine.m}:
\begin{lstlisting}
  w=makeFiniteWeight[{1,0,3}];
  v=makeFiniteWeight[{3,2,1}];
  2*w+v==makeFiniteWeight[{5,2,7}]
  w.v==6
\end{lstlisting}

The use of orthogonal basis in the internal structure of weights allows us to work with weights without complete specification of root system which is useful for study of branching, since roots of the subalgebra can be specified by hand.

\subsection{Root systems}
\label{sec:root-systems}

To specify an algebra of finite or affine type it is enough to fix
its root system. Root systems are represented by two datatypes
\lstinline{finiteRootSystem} and \lstinline{affineRootSystem}. The
latter is an extension of the former. We offer several different
constructors for these datastructures. It is possible to specify
the set of simple roots explicitly, for example to study the
subalgebra $B_2\subset B_4$ we can use the definition
\begin{lstlisting}
  b2b4=makeFiniteRootSystem[ { {1,-1,0,0}, {0,1,0,0} } ]
\end{lstlisting}
There are constructors for root systems of simple finite-dimensional Lie algebras:
\begin{lstlisting}
  b2=makeSimpleRootSystem[B,2]
\end{lstlisting}
We use typographic features of {\it Mathematica} frontend to offer mathematical notation for simple Lie algebras:

@@

Non-twisted affine root systems can be created as affine extensions of finte root systems, e.g.
\begin{lstlisting}
  b2affine = makeAffineExtension[b2]
\end{lstlisting}
In notebook interface this can be written simply as $\hat{B}_2$.

Semisimple Lie algebras can be created as sums of simple:
\begin{lstlisting}[mathescape=true]
  $A_1\oplus A_1$ == finiteRootSystem[2, 2, {finiteWeight[2, {1, 0}], finiteWeight[2, {0, 1}]}]
\end{lstlisting}
%% The problem is here

Predicate \lstinline{rootSystemQ} checks if the object is root systems of finite or affine type.

List of simple roots is the property of root system so it is accessed as \lstinline{rs[simpleRoots]}.

We have implemented several functions to get major properties of root systems. Weyl vector is given by the function \lstinline{rho[rs_?rootSystemQ]}:
\begin{lstlisting}[label=list:1]
  In[1]  =  rho[b2]
  Out[1]  =  finiteWeight[2, {3/2, 1/2}]
\end{lstlisting}
Positive roots can be constructed with \lstinline{positiveRoots[rs_?rootSystemQ]}. For an affine Lie algebra this and related functions return the list up to fixed grade. This grade limit is set through the value of \lstinline{rs[gradeLimit]} which is equal to 10 by default. List of roots (up to \lstinline{gradeLimit}) is returned by \lstinline{roots[rs]}. Cartan matrix and fundamental weights are calculated by the functions \lstinline{cartanMatrix} and \lstinline{fundamentalWeights} correspondingly.

It is possible to specify a weight of a Lie algebra by its Dynkin labels
\begin{lstlisting}
  weight[b2][1,2] == makeFiniteWeight[{2, 1}]
\end{lstlisting}
The function \lstinline{dynkinLabels[rs_?rootSystemQ][wg_?weightQ]} returns Dynkin labels of weight \lstinline{wg} in the root system \lstinline{rs}.

Elements of Weyl group are specified by the set of indices of reflections, so the element  $w=s_{1}s_{2}s_{1}$ of the Weyl group of algebra $B_{2}$ is constructed with \lstinline{weylGroupElement[b2][1,2,1]}. Then it can be applied to the weights:
\begin{lstlisting}
  w = weylGroupElement[b2][1,2,1];
  w @ makeFiniteWeight[{1,0}] == makeFiniteWeight[{-1,0}]
\end{lstlisting}

Computation of lexicographically minimal form \cite{casselman1994machine,casselman1995automata} for Weyl group elements can be conveniently implemented through the use of pattern-matching in {\it Mathematica}. In \cite{KallenShortlex} rewrite rules for simple finite dimensional and affine Lie algebras are presented as {\it Mathematica} patterns. Our presentation of Weyl group elements is compatible with the code of \cite{KallenShortlex}
\begin{lstlisting}[mathescape=true]
  In[1]  = $<<$A3reduce;
           reduce[s[1,2,1,2,1,3,2,1,1]]
  Out[1] = s[2, 3, 2]

@@

\subsection{Formal elements}
\label{sec:formal-elements}

We represent formal characters of representations by special structure \lstinline{formalElement}. This structure is a hash-table implemented with \lstinline{DownValues}. The keys are weights at the exponents and the values are the  corresponding multiplicities. \lstinline[mathescape=true]!makeFormalElement[{$\gamma_{1},\dots,\gamma_{n}$},{$m_{1},\dots,m_{n}$}]! creates data-structure which represents the element $\sum_{i=1}^{n} m_{i} e^{\gamma_{i}}$ of the formal algebra $\mathcal{E}$. The operations in $\mathcal{E}$ are implemented for \lstinline{formalElement} data-type: formal elements can be added, multiplied by a number or by an exponent of a weight. There exists also a multiplication of formal elements but no division.
\begin{lstlisting}[mathescape=true]
  In[1]  = makeFormalElement[{makeFiniteWeight[{1,1}],makeFiniteWeight[{0,0}]},{1,2}] *
             (2 * Exp[makeFiniteWeight[{1,0}]] *
             makeFormalElement[{makeFiniteWeight[{1,1}],makeFiniteWeight[{0,0}]},{1,2}]);
  In[2]  = In[1][weights]
  Out[2] = {finiteWeight[2, {1, 0}], finiteWeight[2, {2, 1}], finiteWeight[2, {3, 2}]}

@@

\subsection{Modules}
\label{sec:modules}

{\bf Affine.m} can be used to study different kinds of modules, i.e. Verma modules, irreducible modules and parabolic Verma modules.  We need datastructure \lstinline{module} to represent generic module of Lie algebra $\gf$. Module properties can be deduced from its set of singular weights using Weyl character formulae \eqref{eq:11},\eqref{eq:12},\eqref{eq:18},\eqref{eq:13}. Set of singular weights can have Weyl symmetry. It can be a symmetry with respect to the Weyl group $W_{\gf}$ or with respect to some subalgebra $W_{\af}$ as in the case of parabolic Verma modules. Then it is possible to study only the main Weyl chamber $C_{\af}$. To use this symmetry generic constructor for \lstinline{module} datastructure accepts several parameters \lstinline{makeModule[rs_?rootSystemQ][singWeights_formalElement,subs_?rootSystemQ|emptyRootSystem[],limit:10}. Here \lstinline{rs} is the root system of the Lie algebra $\gf$,\lstinline{singWeights} is the set of singular weights,  \lstinline{subs} is the root system corresponding to the Weyl group $W_{\af}$ which is the (anti-)symmetry of the set of singular weights. Parameter \lstinline{limit} limits computation for infinite-dimensional modules such as Verma or parabolic Verma.
There are several specialized constructors for different types of highest weight modules:
\begin{lstlisting}[mathescape=true]
vm=makeVermaModule[$B_{2}$][{2,1}];
pm=makeParabolicVermaModule[$B_{2}$][weight[$B_{2}$][2,1],{1}];
im=makeIrreducibleModule[$B_{2}$][2,1];
GraphicsRow[textPlot/@{im,vm,pm}]
$\includegraphics[width=130mm]{figures/irrep-verma-pverma}$
\end{lstlisting}

>>>>

Бесконечномерные алгебры Ли гораздо сложнее исследовать и число имеющихся программ гораздо меньше. 
Однако структура аффинных алгебр Ли позволяет адаптировать к ним вычислительные алгоритмы, предложенные для конечномерных алгебр Ли \cite{Fuchs:1996dd}, \cite{gannon2001algorithms}, \cite{kass1990ala}. В книге \cite{kass1990ala}, изданной в 1990 году, приведены таблицы кратностей весов неприводимых представлений и другие вычисленные характеристики аффинных алгебр и их модулей. Однако нам на данный момент не известны пакеты для популярных систем компьютерной алгебры, которые можно было бы использовать для воспроизведения и расширения этих результатов.

Чтобы исправить этот недостаток, нами был создан пакет {\bf Affine.m} для популярной системы  {\it Mathematica}. Возможности и ограничения этого пакета мы и описываем в этой главе. Кроме того, мы приводим некоторые примеры, связанные с физическими приложениями.

Необходимые сведения из теории представлений приведены в главе \ref{cha:affine-lie-algebras}. Здесь мы начинаем с описания структур данных пакета  {\bf Affine.m}, использующихся для описания различных объектов теории представлений  (раздел \ref{sec:core-datastructures}), обсуждаем алгоритмы  (раздел \ref{sec:comp-algor}) и приводим примеры  (раздел \ref{sec:examples}).

\section{Структуры данных}
\label{sec:core-datastructures} 
Хотя {\it Mathematica} -- нетипизированный язык программирования, в нем можно создавать структурированные объекты и проверять их типы используя сопоставление с шаблоном (pattern matching) \cite{shifrinmathematica}, \cite{maeder2000computer}.
\subsection{Веса}
\label{sec:weights}

Веса представляются двумя структурами данных: \lstinline{finiteWeight} для конечномерных алгебр Ли и \lstinline{affineWeight} для аффинных алгебр Ли.

Внутренняя структура веса конечномерной алгебры Ли представляет собой  \lstinline{List} с заголовком \lstinline{finiteWeight}, его компоненты -- это координаты веса в ортогональном базисе выбранном согласно Бурбаки \cite{bourbaki2002lie}.
Аффинный вес представляет собой расширение конечномерного веса с добавлением уровня и грейда.

В пакете {\bf Affine.m} содержится набор функций для работы с весами конечномерных и аффинных алгебр Ли. Полный список доступен во встроенной справке пакета. Важнейшие из них -- это определение операций сложения, умножения на число и скалярного произведения (билинейной формы) для весов. В результате можно использовать традиционные обозначения при работе с  {\bf Affine.m}:
\begin{lstlisting}
  w=makeFiniteWeight[{1,0,3}];
  v=makeFiniteWeight[{3,2,1}];
  2*w+v==makeFiniteWeight[{5,2,7}]
  w.v==6
\end{lstlisting}

Использование ортогонального базиса во внутренней структуре весов позволяет нам работать с весами без полного задания корневой системы. Это полезно при изучении редукции на подалгебры, так как корни подалгебры можно задать вручную, путем указания их координат в ортогональном базисе.

\subsection{Корневые системы}
\label{sec:root-systems}

Чтобы задать конечномерную или аффинную алгебру достаточно определить ее корневую систему. Корневые системы представляются двумя структурами данных \lstinline{finiteRootSystem} и \lstinline{affineRootSystem}. Вторая структура является расширением первой. Мы предлагаем несколько конструкторов для этих структур данных. Во-первых, можно определить простое корни явно, например, при изучении вложения $B_2\subset B_4$ можно использовать определение
\begin{lstlisting}
  b2b4=makeFiniteRootSystem[ { {1,-1,0,0}, {0,1,0,0} } ]
\end{lstlisting}
Для корневых систем простых конечномерных алгебр Ли есть специальные конструкторы:
\begin{lstlisting}
  b2=makeSimpleRootSystem[B,2]
\end{lstlisting}
В графическом интерфейсе {\it Mathematica} можно использовать традиционные математические обозначения для простых алгебр Ли:

@@

Нескрученные аффинные корневые системы могут быть созданы как аффинные расширения корневых систем конечномерных алгебр Ли, например:
\begin{lstlisting}
  b2affine = makeAffineExtension[b2]
\end{lstlisting}
В графическом интерфейсе это же определение можно ввести просто как $\hat{B}_2$.

Полупростые алгебры Ли представляют собой прямые суммы простых:
\begin{lstlisting}[mathescape=true]
  $A_1\oplus A_1$ == finiteRootSystem[2, 2, {finiteWeight[2, {1, 0}], finiteWeight[2, {0, 1}]}]
\end{lstlisting}
%% The problem is here

Предикат \lstinline{rootSystemQ} проверяет, является ли объект корневой системой конечномерной или аффинной алгебры Ли.

Список простых корней -- это определяющее свойство корневой системы, поэтому он доступен как \lstinline{rs[simpleRoots]}.

В пакете реализовано несколько функций для основных характеристик корневой системы. Вектор Вейля дается функцией
\lstinline{rho[rs_?rootSystemQ]}:
\begin{lstlisting}[label=list:1]
  In[1]  =  rho[b2]
  Out[1]  =  finiteWeight[2, {3/2, 1/2}]
\end{lstlisting}
Положительные корни строятся функцией  \lstinline{positiveRoots[rs_?rootSystemQ]}. Для аффинной алгебры Ли эта и аналогичные функции возвращают список до некоторого фиксированного грейда. Это максимальное значение грейда задается как свойство \lstinline{rs[gradeLimit]} и по умолчанию равняется 10. Список корней  (вплоть до грейда \lstinline{gradeLimit}) строится функцией \lstinline{roots[rs]}. Матрица Картана и фундаментальные веса вычисляются функциями \lstinline{cartanMatrix} и \lstinline{fundamentalWeights} соответственно.

Вес алгебры Ли можно задать его индексами Дынкина
\begin{lstlisting}
  weight[b2][1,2] == makeFiniteWeight[{2, 1}]
\end{lstlisting}
Функция \lstinline{dynkinLabels[rs_?rootSystemQ][wg_?weightQ]} возвращает индексы Дынкина веса \lstinline{wg} по отношению к корневой системе \lstinline{rs}.

Элементы группы Вейля задаются номерами элементарных отражение, так что элемент  $w=s_{1}s_{2}s_{1}$ группы Вейля алгебры Ли $B_{2}$ строится вызовом функции \lstinline{weylGroupElement[b2][1,2,1]}. После этого его можно применить к весам:
\begin{lstlisting}
  w = weylGroupElement[b2][1,2,1];
  w @ makeFiniteWeight[{1,0}] == makeFiniteWeight[{-1,0}]
\end{lstlisting}

Вычисление лексикографически минимальной формы \cite{casselman1994machine,casselman1995automata} для элементов группы Вейля можно реализовать при помощи шаблонов в  {\it Mathematica}. В работе \cite{KallenShortlex} представлены правила подстановки на языке {\it Mathematica} для такого вычисления в случае простых конечномерных и аффинных алгебр Ли. Наше представление для элементов группы Вейля совместимо с кодом \cite{KallenShortlex}:
\begin{lstlisting}[mathescape=true]
  In[1]  = $<<$A3reduce;
           reduce[s[1,2,1,2,1,3,2,1,1]]
  Out[1] = s[2, 3, 2]

@@

\subsection{Формальные элементы}
\label{sec:formal-elements}

Формальные элементы представляются специальной структурой данных \lstinline{formalElement}. Эта структура данных представляет собой хэш-таблицу, реализованную через механизм \lstinline{DownValues}. Ключами в таблице выступают веса, стоящие в экспонентах формального элемента, а значениями -- соответствующие кратности.  Вызов \lstinline[mathescape=true]!makeFormalElement[{$\gamma_{1},\dots,\gamma_{n}$},{$m_{1},\dots,m_{n}$}]! создает структуру данных, представляющую элемент $\sum_{i=1}^{n} m_{i} e^{\gamma_{i}}$ формальной алгебры $\mathcal{E}$. Операции в алгебре  $\mathcal{E}$ реализованы для структуры \lstinline{formalElement}: формальные элементы можно складывать, умножать на число или экспоненту веса. Кроме того, есть умножение формальных элементов, но нет операции деления.
\begin{lstlisting}[mathescape=true]
  In[1]  = makeFormalElement[{makeFiniteWeight[{1,1}],makeFiniteWeight[{0,0}]},{1,2}] *
             (2 * Exp[makeFiniteWeight[{1,0}]] *
             makeFormalElement[{makeFiniteWeight[{1,1}],makeFiniteWeight[{0,0}]},{1,2}]);
  In[2]  = In[1][weights]
  Out[2] = {finiteWeight[2, {1, 0}], finiteWeight[2, {2, 1}], finiteWeight[2, {3, 2}]}

@@

\subsection{Модули}
\label{sec:modules}

{\bf Affine.m} можно использовать для изучения различных видов модулей, например, модулей Верма, неприводимых модулей и параболических модулей Верма. Для представления произвольного модуля используется структура данных  \lstinline{module}.

Module properties can be deduced from its set of singular weights using Weyl character formulae \eqref{eq:11},\eqref{eq:12},\eqref{eq:18},\eqref{eq:13}. Set of singular weights can have Weyl symmetry. It can be a symmetry with respect to the Weyl group $W_{\gf}$ or with respect to some subalgebra $W_{\af}$ as in the case of parabolic Verma modules. Then it is possible to study only the main Weyl chamber $C_{\af}$. To use this symmetry generic constructor for \lstinline{module} datastructure accepts several parameters \lstinline{makeModule[rs_?rootSystemQ][singWeights_formalElement,subs_?rootSystemQ|emptyRootSystem[],limit:10}. Here \lstinline{rs} is the root system of the Lie algebra $\gf$,\lstinline{singWeights} is the set of singular weights,  \lstinline{subs} is the root system corresponding to the Weyl group $W_{\af}$ which is the (anti-)symmetry of the set of singular weights. Parameter \lstinline{limit} limits computation for infinite-dimensional modules such as Verma or parabolic Verma.
There are several specialized constructors for different types of highest weight modules:
\begin{lstlisting}[mathescape=true]
vm=makeVermaModule[$B_{2}$][{2,1}];
pm=makeParabolicVermaModule[$B_{2}$][weight[$B_{2}$][2,1],{1}];
im=makeIrreducibleModule[$B_{2}$][2,1];
GraphicsRow[textPlot/@{im,vm,pm}]
$\includegraphics[width=130mm]{figures/irrep-verma-pverma}$
\end{lstlisting}
